---
title: "Random Forest"
author: "Anthony Staines"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}

library(tidyverse)
library(randomForest)
library(ranger)

knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
set.seed(412789)
```

# Training data

Start with the Ames housing data. This is a supervised learning example dataset, where the goal is to predict the sales price. See De Cock, D. (2011). "Ames, Iowa: Alternative to the Boston Housing Data as an End of Semester Regression Project," Journal of Statistics Education, Volume 19, Number 3.
Note that Sale_Price is directly observed, without error.

```{r}
AMES <- AmesHousing::make_ames()
```

We also use the Atttrition data set, which is employee attrition information originally provided by the IBM Watson Analytics Lab. 
```{r}

data("attrition", package = "modeldata")
glimpse(attrition)
Attrition <- attrition
rm(attrition)

```
